\documentclass{article}
\usepackage{graphicx} % Required for inserting image
\begin{document}
\textbf{{Machine Learning Theory Questions}}
\section*{Supervised Machine Learning}
\begin{itemize}
    \item[(a)] Define supervised learning. Give two examples of its applications.
    \item[(b)] What is the difference between regression and classification?
    \item[(c)] Why do we use a cost function in linear regression? How does it relate to optimization?
    \item[(d)] Describe how logistic regression can be used for binary classification.
    \item[(e)] What is the role of the sigmoid function in logistic regression?
\section*{Advanced Learning Algorithms}
    \item[(a)] What problem does regularization solve in machine learning? How does it work?
    \item[(b)] What is the difference between bias and variance in a learning algorithm?
    \item[(c)] Define precision and recall. Why are they important in model evaluation?
    \item[(d)] Explain the basic structure of an artificial neural network.
    \item[(e)] What is the vanishing gradient problem in deep neural networks?
\section*{Unsupervised Learning}

    \item[(a)] Define unsupervised learning. How does it differ from supervised learning?
    \item[(b)] What is the main objective of K-Means clustering?
    \item[(c)] Describe one real-world application of clustering algorithms.
    \item[(d)] What is PCA (Principal Component Analysis), and why is it used?
    \item[(e)] How can we choose the number of clusters \( K \) in K-Means?
\section*{Anomaly Detection vs Supervised Learning}
\textbf{(a)} Explain the key differences between supervised learning and anomaly detection in terms of:
    \item Training data
    \item Number of positive examples
    \item Use cases
\end{itemize}
\vspace{0.5cm}

\textbf{(b)} You are building a system to detect fraudulent credit card transactions. Would you choose a supervised learning algorithm or an anomaly detection algorithm? Justify your choice.
\vspace{0.5cm}

\textbf{(c)} Fill in the table below to compare the two approaches:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Aspect} & \textbf{Supervised Learning} & \textbf{Anomaly Detection} \\
\hline
Label availability & & \\
\hline
Number of positive examples & & \\
\hline
Best used when & & \\
\hline
Example application & & \\
\hline
\end{tabular}
\end{center}

\end{document}
